Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-22 paper.tex
Add fig S2 hardware details

We used bcftools version 1.18, Savvy 2.1.0, Genozip 5.0.26,
vcf2zarr 0.0.9, and Zarr-Python 2.17.2. 
All tools used default settings, 
unless otherwise stated.
All simulation-based benchmarks were performed on a 
dual CPU (Intel Xeon E5-2680 v2)
server with 256GiB of RAM running Debian GNU/Linux 11.
To ensure that the true effects of having data distributed over a large 
number of files were reported, benchmarks for Zarr and Savvy were 
performed on a cold disk cache by running 
\texttt{echo 3 | sudo tee /proc/sys/vm/drop\_caches} before each run.
The I/O subsystem used is based on a RAID 5 of 12 SATA hard drives.
For the CPU time benchmarks we measure the sum of the total user and
system times required to execute the full command (as reported by GNU
\texttt{time}) as well as elapsed wall-clock time. Total CPU
time is shown as a solid line, with wall-clock time as a dashed line 
of the same colour. In the case of pipelines, where some processing 
is conducted concurrently wall-clock time can be less than total
CPU (e.g.\ genozip in Fig~\ref{fig-whole-matrix-compute}). 
When I/0 costs are significant, wall-clock time can be greater 
than total CPU (e.g.\ Zarr and Savvy in Fig~\ref{fig-subset-matrix-compute}).
Each tool was instructed to use one thread, where the options
were provided. 
Where possible in pipelines we use uncompressed BCF
output (\texttt{-Ou}) to make processing 
more efficient~\citep{danecek2021twelve}.
We do not use BCF output in genozip because it is not supported
directly.

@@

\begin{figure}[h]
\includegraphics{figures/whole-matrix-compute-zarr-versions}
\caption{Whole-matrix computation performance using different Zarr 
implementations. 
Total CPU time required to run the af-dist calculation
the data in Figure~\ref{fig-data-storage} using different Zarr implementations.
Elapsed time is also reported (dotted line). 
These benchmarks were run on [TODO insert hardware details].
Note the difference between the
time reported here and in Fig~\ref{fig-whole-matrix-compute} for Zarr-Python
is due to different hardware platforms.
\label{fig-whole-matrix-compute-zarr-versions}}
\end{figure}

>>>>

We used bcftools version 1.18, Savvy 2.1.0, Genozip 5.0.26,
vcf2zarr 0.0.9, and Zarr-Python 2.17.2. 
All tools used default settings, 
unless otherwise stated.
All simulation-based benchmarks (excluding Fig~\ref{fig-whole-matrix-compute-zarr-versions}) were performed on a 
dual CPU (Intel Xeon E5-2680 v2)
server with 256GiB of RAM running Debian GNU/Linux 11.
To ensure that the true effects of having data distributed over a large 
number of files were reported, benchmarks for Zarr and Savvy were 
performed on a cold disk cache by running 
\texttt{echo 3 | sudo tee /proc/sys/vm/drop\_caches} before each run.
The I/O subsystem used is based on a RAID 5 of 12 SATA hard drives.
For the CPU time benchmarks we measure the sum of the total user and
system times required to execute the full command (as reported by GNU
\texttt{time}) as well as elapsed wall-clock time. Total CPU
time is shown as a solid line, with wall-clock time as a dashed line 
of the same colour. In the case of pipelines, where some processing 
is conducted concurrently wall-clock time can be less than total
CPU (e.g.\ genozip in Fig~\ref{fig-whole-matrix-compute}). 
When I/0 costs are significant, wall-clock time can be greater 
than total CPU (e.g.\ Zarr and Savvy in Fig~\ref{fig-subset-matrix-compute}).
Each tool was instructed to use one thread, where the options
were provided. 
Where possible in pipelines we use uncompressed BCF
output (\texttt{-Ou}) to make processing 
more efficient~\citep{danecek2021twelve}.
We do not use BCF output in genozip because it is not supported
directly.

@@

\begin{figure}[h]
\includegraphics{figures/whole-matrix-compute-zarr-versions}
\caption{Whole-matrix computation performance using different Zarr 
implementations. 
Total CPU time required to run the af-dist calculation
the data in Figure~\ref{fig-data-storage} using different Zarr implementations.
Elapsed time is also reported (dotted line). 
These benchmarks were run on an 8-core CPU (Intel i7-9700) with 32GiB RAM running
Linux Mint 21.3 with data on an NVMe SSD.
Note the difference between the
time reported here and in Fig~\ref{fig-whole-matrix-compute} for Zarr-Python
is due to different hardware platforms.
\label{fig-whole-matrix-compute-zarr-versions}}
\end{figure}
